Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2018-02-10 paper.tex

%Genomic evidence points to a history of wild-to-crop introgression during dispersal of barley beyond their centers of domestication, but for the most part, the identity of the alleles critical to these adaptations have yet to be uncovered.
%One exception comes from the authors of \cite{dai2012tibet}, who hypothesize that alleles for cold stress tolerance and spring habit \gmj{"spring habit" are the words from the paper, but they're not defined.  Does this mean that it grows more quickly in the spring?} from wild Tibetan 2-rowed barley allowed the barley to spread to the cold northern regions of Europe.
%Wild-domesticate breeding experiments have shown that wild barleys have alleles for several important fitness and agronomic phenotypes, including powdery mildew resistance, brittleness, flowering time, plant height, lodging, and yield \cite{dreiseitl2017heterogeneity,von2006ab,handley1994chromosome}, and these individually or in combination may have been facilitators of the dispersal of domesticated barley.

@@

The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
Certain genetic and archaeobotanical evidence points toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China 7000 BP and the Indian Ganges plain 4000 BP, respectively \cite{fuller2010consilience}, with possibly another domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012}.
Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of variety \emph{aromatic}.
A third analysis \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose via a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which became the progenitors to the other Asian rice varieties.
 %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."

%The former first paragraph, replaced from above:
%The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
%Interspecific hybridization between wild and domesticated rices both ancient and recent has produced a phylogeny that defies clear and simple models.
%On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
%Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
%A recent investigation utilizing SNP data from wild and domesticated rice accesssions pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from rufipogon, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
%Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia. %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."

%The former first paragraph, replaced from above:
%The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
%A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
%\cite{vaughan2008evolving} support this view, noting that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles.
%On the other hand, genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
%An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of \emph{japonica} in China followed by formation of the \emph{indica} group through hybridization of \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
%Adaptive introgression may in fact be the signal interpreted as independent domestications.

The findings of \cite{choi2018multiple} run in parallel with a hypothesis posited by Vaughan and colleagues in 2008 \cite{vaughan2008evolving}.
%Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
According to this hypothesis, the non-shattering allele (\emph{sh1}) arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like) in the region.
This would have facilitated the subsequent spread of other domestication alleles among those populations.
As these domesticated populations then spread into new environments, they would have received introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
Both Vaughan and Choi support the role of introgression between domesticated Asian rice and locally-adapted wild \emph{rufipogon} populations in the development and dispersal of Asian rice into new environments.

Despite many conditions favorable to their prevalence, to date, individual examples of adaptive alleles naturally introgressing into domesticated Asian rice are not clear (though \cite{zhao2010genomic} emphasizes the necessity of research on this topic).
Contemporary populations of wild rices show evidence of this history of local adaptation.
Wild \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
Domesticated varieties have likewise adapted to local environments.
The \emph{japonicas} are generally more adapted to upland temperate regions (and are split into subgroups adapted to more temperate or more tropical conditions), the \emph{indicas} are best suited for tropical lowland environments, \emph{aus} matures quickly and tolerates drought stress, and two deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river \cite{vaughan2008evolving, garris2005genetic}.
Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
However, as with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
Still, barriers to crop-to-wild introgression may have been less severe during early domestication as the inbreeding reproductive system of domesticated rice would not have been as firmly established \cite{vaughan2008evolving}.
Furthermore, the contemporary distribution of wild rice does not capture the full range and diversity of wild rice during early domestication and range expansion of Asian rice.

>>>>

@@

The details of Asian rice (\emph{Oryza sativa}) domestication are still debated.
Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China 7000 BP and the Indian Ganges plain 4000 BP, respectively \cite{fuller2010consilience}, with a potential third domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012}.
Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between a range-wide sample of wild and domesticated rice, finding \emph{japonica} was domesticated near the Pearl River in Guangxi province, China, and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of \emph{aromatic} rice.
A recent third analysis by Choi and co-authors \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose via a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which thereby became the progenitors of other Asian rice varieties.

The findings of \cite{choi2018multiple} run in parallel with a hypothesis posited by Vaughan and colleagues \cite{vaughan2008evolving}.
According to this hypothesis, domestication alleles arose in a single cultivated rice population and subsequently introgressed into diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like) in the region.
As these fully domesticated populations spread into new environments, they potentially received introgression from locally-adapted wild relatives, retaining alleles that improved fitness.

%Despite many conditions favorable to their prevalence, to date, individual examples of adaptive alleles naturally introgressing into domesticated Asian rice are not clear (though \cite{zhao2010genomic} emphasizes the necessity of research on this topic).
%Contemporary populations of wild rices show evidence of this history of local adaptation.
%Wild \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
%Domesticated varieties have likewise adapted to local environments.
%The \emph{japonicas} are generally more adapted to upland temperate regions (and are split into subgroups adapted to more temperate or more tropical conditions), the \emph{indicas} are best suited for tropical lowland environments, \emph{aus} matures quickly and tolerates drought stress, and two deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river \cite{vaughan2008evolving, garris2005genetic}.
%Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
%However, as with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
%This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
%Still, barriers to crop-to-wild introgression may have been less severe during early domestication as the inbreeding reproductive system of domesticated rice would not have been as firmly established \cite{vaughan2008evolving}.
%Furthermore, the contemporary distribution of wild rice does not capture the full range and diversity of wild rice during early domestication and range expansion of Asian rice.
